---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hubverse

<!-- badges: start -->
<!-- badges: end -->

## Overview

A collection of packages that enables collaborative modeling exercises through a unified framework for aggregating, visualizing, and evaluating forecasts. This package is designed to make it easy to install and load multiple 'hubverse' packages in a single step.

Learn more at <https://hubverse.io/en/latest/index.html>.

## Installation

You can install the development version of hubverse from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Infectious-Disease-Modeling-Hubs/hubverse")
```

## Usage

`library(hubverse)` loads the core hubverse packages:

- [hubData](https://infectious-disease-modeling-hubs.github.io/hubData/) for tools to access and work with data in a modeling hub
- [hubEnsembles](https://infectious-disease-modeling-hubs.github.io/hubEnsembles/) for ensemble methods to combine model outputs
- [hubVis](https://infectious-disease-modeling-hubs.github.io/hubVis/) for plotting model outputs
- [hubUtils](https://infectious-disease-modeling-hubs.github.io/hubUtils/) for lightweight utility functions
- [hubExamples](https://infectious-disease-modeling-hubs.github.io/hubExamples/) for example data in the hubverse format
- [hubEvals](https://infectious-disease-modeling-hubs.github.io/hubEvals/) for evaluating model outputs
- [hubValidations](https://infectious-disease-modeling-hubs.github.io/hubValidations/) for validating modeling hubs and their model output
- [hubAdmin](https://infectious-disease-modeling-hubs.github.io/hubAdmin/) for administrating a modeling hub
- [hubCI](https://infectious-disease-modeling-hubs.github.io/hubCI/) for continuous integration tools
- [hubDevs](https://infectious-disease-modeling-hubs.github.io/hubDevs/) for creating and standardizing hubverse packages

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(hubverse)
## basic example code
```

***

## Code of Conduct

Please note that the hubEvals package is released with a [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.

## Contributing

Interested in contributing back to the open-source Hubverse project?
Learn more about how to [get involved in the Hubverse Community](https://hubdocs.readthedocs.io/en/latest/overview/contribute.html)
